A Technique for Inter-Catalogue Recommendations

### 2014-06-10 paper2/paper.tex

First, we scrape all of the clothing products from Macy's website at 
www.macys.com. For each product we get its image, what category of clothing it
is in, its description, and a list of other products recommended for it.

Next, we simulate the catalogues of two different retailers by partitioning all
of the products into two disjoint collections of products. To make this
partition realistic, we ensure that no brand has products in both catalogue. We
remove all recommendations between products from one partition to the other,
thus creating two entirely disjoint and unconnected catalogues.

Now, for each product in each catalogue, we choose which products to recommend
in the other catalogue, using the recommendation method to be tested. To make
this problem slightly simpler, we restrict recommendations to be within the same
category. In this way, connecting up entire catalogues becomes a series of
smaller tasks of connecting the products in each category.

We still have all of the true inter-catalogue recommendations from the
unpartitioned products. Although we don't use this information for constructing
recommendations between the partitioned graphs, we can use the original
recommendations as a gold-standard to test the construct ones. Using these
gold-standard recommendations, we can evaluate the precision and recall of the
predicted inter-catalogue recommendations.

For the Macy's catalogue, each product has four recommendations. This means that
on average, there are only two gold-standard inter-catalogue recommendations
from each product. However, since there potentially are other good
recommendations besides these two, we create another metric that expands our
notion of a correct recommendation. Here we look at recommendations of
recommendations. For example, if Product B is recommended for Product A and
Product C is recommended for Product B, we say that Product C is a good
recommendation for Product A. This leads us to the notion of 2-precision and
2-recall.

@@

2-precision is defined as the percentage of predicted edges that are within two
steps in the original, unpartitioned recommender graph. In the example above,
Product A and Product C are within two steps in the original graph, thus an edge
between A and C is considered correct and increases precision. 2-recall is
defined analogously. Here, we consider the new recommender graph consisting of
the two partitioned recommender graphs connected by the predicted edges.
2-recall is then the percentage of gold-standard edges that are within two steps
in the predicted recommender graph.

@@

\subsection*{Content Based Similarity}
A simple approach for this problem is to recommend products that have similar
content. The products have both a short and long description which we can use to
construct a content-based similarity function. To do this, we represent each
product as a vector of tf-idf values for each word in its descriptions. By using
tf-idf, we reduce the relevance of frequent words and boost that of rare ones.
The vectors are the length of the dictionary, with zeros for all absent words
from the descriptions. To determine the content-based similarity between two
products, we simply take the dot product of the two tf-idf vectors.

@@

While this is a quadric-time operation, it can be sped up using matrix
multiplication. Furthermore, the division into different categories reduces the
impact of the quadratic effect. For instance, Macy's entire catalogue consists
of over one million products, making a quadratic time algorithm infeasible.
However, the average category has 2000 products, which is significantly more
manageable.

\subsection*{Leveraging Recommendations}
The simple content-based recommender is limited for a few reasons. It attempts
to capture similarity between products by looking exclusively at the words used
to describe them. One limitation here is that different brands often use
different vocabularies to describe the same thing, such as ``off-white'' versus
``ivory.'' Another issue is that the best recommendation for a product is often
more subtle than simply the most similar other product. For instance, a striped
black-and-white button up shirt may be an excellent recommendation for a blue
checkered button-up, even though the two have ostensibly different patterns and
colors.

@@

where...

Assuming the recommendations are good within each of the catalogues, this
approach can help both of the issues brought up with the simple content-based
recommender. Since a neighborhood has many products, each describing themselves
with potentially different vocabularies, there is a higher likelihood that the
neighborhood contains multiple synonyms and thus decreases the potential of
missing due to different vocabularies. Furthermore, the recommender graphs
capture some of the user behavior not expressed in product descriptions. Looking
back on the dress-shirts example, the neighborhood of a striped shirt may
include checkered shirts and vise-versa. In this case, the two neighborhoods
would share the words ``striped'' and ``checkered'' and thus the two previously
dissimilar shirts would be considered similar.

\subsection*{Latent Concepts}
Although the NeighborhoodSimilarity method alleviates some of the issues of
content-similarity, it is still fundamentally based on the descriptions of
products. Ideally, we would like to recommend products that are similar in more
fundamental ways than their descriptions. If a dress is elegant and modern, we
want to recommend another dress that is elegant and modern, even if those words
are not explicitly in the descriptions of either. These types of latent concepts
can be intangible and difficult to define. However, if we assume that the given
recommendations are based on these concepts, we can attempt to learn the latent
concepts by reverse-engineering the graphs.

One approach to learning latent concepts is Latent Semantic Indexing or LSI. In
traditional LSI, a dataset consists of documents, each containing terms.
Singular Value Decomposition (SVD) is applied to the term-document matrix and
the resulting three matrices are truncated to the number of latent concepts
desired. The resulting left matrix is a term-concept matrix, where each row is a
term represented in concept space.

@@

To simulate user sessions, we use a random walk technique on the graph. We
assume that the edges of a recommender graph indeed connect products that a user
would co-view. Thus traversing the edges of a graph is a reasonable simulation
of a user session. For details on these random walks, see the section titled
Random Walks below.

With the simulated user sessions, we can implement LSI on each of the catalogues
to learn the latent concepts for each product. We want to find the products from
one catalogue that are most similar in latent concept space to products in the
other catalogue. However, since the user sessions are disjoint, we must learn
different latent spaces for each catalogue. The problem is therefore reduced to
finding a mapping from one concept space to the other. If this can be done, we
can map the concepts of each product into concepts of the other catalogue and
then find its nearest neighbors in that catalogue's concept space.

>>>>

First, we scrape all of the clothing products from Macy's website at 
www.macys.com. For each product we save its image, what category of clothing it
is in, its description, and a list of other products recommended for it.

Next, we partition all of the Macy's products into two disjoint sets to simulate 
the catalogues of two different retailers. To make this partition realistic, 
we ensure that no brand has products in both catalogue. We remove all 
recommendations between products from one partition to the other, thus creating 
two entirely disjoint and unconnected catalogues.

Now, we can use this partition to test a method for constructing recommendation.
For each product in each of the two catalogues, we use our recommender to choose 
which products to recommend in the other catalogue. To make this problem slightly
simpler, we restrict recommendations to be within the same category, such as
Women's Dresses or Men's Shoes. In this way, connecting up entire catalogues 
becomes a series of smaller tasks of connecting the products in each category.

For this experiment, we have access to all of the true inter-catalogue 
recommendations from the unpartitioned Macy's catalogue. Although we don't
use this information for constructing recommendations between the partitioned 
graphs, we can use the original recommendations as a gold-standard to test the 
construct ones. Using these gold-standard recommendations, we can evaluate 
the precision and recall of the predicted inter-catalogue recommendations.

For the Macy's catalogue, each product has four recommendations. This means that
on average, there are only two gold-standard inter-catalogue recommendations
from each product. Since there potentially are other good
recommendations besides these two, we create another metric that expands our
notion of a correct recommendation. Here we extend the definition of a good
connection to include recommendations of recommendations in the original
catalgoue. For example, if Product B is recommended for Product A and 
Product C is recommended for Product B, we say that Product C is still a good 
recommendation for Product A. This leads us to the notions of 2-precision and 
2-recall.

@@

2-precision is defined as the percentage of predicted edges that are within two
steps in the original, unpartitioned recommender graph. In the example above,
Product A and Product C are within two steps in the original graph, thus an edge
between A and C is considered correct and increases precision. 2-recall is
defined analogously. Here, we consider the new recommender graph, consisting of
the two partitioned recommender graphs connected by the predicted
recommendations. 2-recall is then the percentage of gold-standard edges that 
are within two steps in the predicted recommender graph.

@@

\subsection*{Content Based Similarity}
A simple approach for this problem is to recommend products that have similar
content. The products have both a short and long description which we can use to
construct a textual content-based similarity function. To do this, we represent each
product as a vector of tf-idf values for each word in its descriptions. By using
tf-idf, we reduce the relevance of frequent words and boost that of rare ones,
thus emphasizing the more words that are more descriptive of a product versus
all the other ones in the catalogue. The vector representations are the length 
of the dictionary, with zeros for all words not in the descriptions. To determine
the content-based similarity between two products, we simply take the dot 
product of the two tf-idf vectors.

Mathematical definition of content-based similairty...

@@

While this is a quadric-time operation, it can be sped up using matrix
multiplication. Furthermore, the division into different categories reduces the
impact of the quadratic effect. For instance, Macy's entire catalogue consists
of over ONE MILLION products, making a quadratic time algorithm infeasible.
However, the average category has 2000 products, which is significantly more
manageable.

\subsection*{Leveraging Recommendations}
The simple content-based recommender is limited for a few reasons. It attempts
to capture similarity between products by looking exclusively at the words used
to describe them. One limitation here is that different brands often use
different vocabularies to describe the same thing, such as ``off-white'' versus
``ivory.'' Another issue is that the best recommendation for a product is often
more subtle than simply the most similar other product. For instance, a striped
black-and-white button-up shirt may be an excellent recommendation for a blue
checkered button-up, even though the two have ostensibly different patterns and
colors.

@@

NeighborhoodSimilarity definition...

Assuming the recommendations are good within each of the catalogues, this
approach can help both of the issues brought up with the simple content-based
recommender. Since a neighborhood has many products, each describing themselves
with potentially different vocabularies, there is a higher likelihood that the
neighborhood contains multiple synonyms and thus decreases the potential of
missing due to different vocabularies. Furthermore, the recommender graphs
capture some of the user behavior not expressed in product descriptions.
Re-examining the button-up shirts example, the neighborhood of a striped shirt may
include checkered shirts and vise-versa. In this case, the two neighborhoods
would share the words ``striped'' and ``checkered'' and thus the two previously
dissimilar shirts would now be considered similar.

\subsection*{Latent Concepts}
Although the NeighborhoodSimilarity method alleviates some of the issues of
content-similarity, it is still fundamentally based on the descriptions of
products. Ideally, we would like to recommend products that are similar in more
fundamental ways than their descriptions. If a dress is elegant and modern, we
want to recommend another dress that is elegant and modern, even if those words
are not explicitly in the descriptions of either. These types of latent concepts
can be intangible and difficult to define. However, if we assume that the given
recommendations are truly based on these underlying concepts, we can attempt to 
learn the latent concepts by analyzing the graphs.

One approach to learning latent concepts is Latent Semantic Indexing or LSI. In
traditional LSI, used for Natural Language Processing, concepts are learned over 
a dataset consists of documents, each containing terms. Singular Value 
Decomposition (SVD) is applied to the term-document matrix and the resulting 
three matrices are truncated to the number of latent concepts desired. The 
resulting left matrix is a term-concept matrix, where each row is a term 
represented in concept space.

@@

To simulate user sessions, we use a random walk technique on the graph. We
assume that the edges of a recommender graph indeed connect products that a user
would co-view. Thus traversing the edges of a graph is a reasonable simulation
of a user session. For details on these random walks, see the section titled
Random Walks in the Implementation section below.

With the simulated user sessions, we can implement LSI on each of the catalogues
and learn the latent concepts for each product. We want to find the products from
one catalogue that are most similar in latent concept space to products in the
other catalogue. However, since the user sessions (random walks) include
products either from one catalogue or the other, we learn different latent 
spaces for each catalogue. The problem is therefore reduced to finding a mapping
from one concept space to the other. If this can be accomplished, we
can map the concepts of each product into concepts of the other catalogue. This
creates a concept representation in the other catalouge's concept space, meaning
that we can then find nearest neighbors in that catalogue's concept space.
